---
title: "DYNAMIC CLASSROOM MANAGEMENT SYSTEM"
output: html_notebook
---


```{r}
set.seed(112233)

n <- 1000

# 1. Engagement score (0–10)
engagement <- round(runif(n, 0, 10), 1)

# 2. Classroom participation (0–100%)
participation <- round(runif(n, 0, 100), 1)

# 3. Attendance (50–100%)
attendance <- round(runif(n, 50, 100), 1)

# 4. Seating arrangement group size (1 = solo, 3 = small group, 6 = large group)
seating_group <- sample(c(1, 3, 6), n, replace = TRUE)

# 5. 
class_duration <- sample(c("Short", "Medium", "Long"), n, replace = TRUE)

# 6. Teaching method (Lecture / Discussion / Interactive / Digital)
teaching_method <- sample(c("Discussion", "Interactive"),
                           n, replace = TRUE)

# Logistic model for classroom outcome
linear_pred <- -10 +
  0.45 * engagement +
  0.03 * participation +
  0.07 * attendance

prob_success <- exp(linear_pred) / (1 + exp(linear_pred))

# 7. Outcome: 1 = No Intervention, 0 = Intervention Needed
class_outcome <- rbinom(n, 1, prob_success)

data_classroom <- data.frame(
  engagement,
  participation,
  attendance,
  seating_group,
  class_duration,
  teaching_method,
  class_outcome = factor(class_outcome,
                         levels = c(0,1),
                         labels = c("Intervention Needed", "No Intervention"))
)

data_classroom

```
```{r}
write.csv(data_classroom, "sms_data_classroom.csv", row.names = FALSE)
```

LOAD DATA
```{r}
data<- read.csv("/Users/sitikhairunnisajannahroslan/Downloads/sms_data_classroom.csv")
data
```
PREPROCESSING DATA
```{r}
library(dplyr)
library(caret)
library(ggplot2)
library(moments)
library(car)
```
Change the data type
```{r}
data$seating_group <- as.factor(data$seating_group)
data$teaching_method <-as.factor(data$teaching_method)
data$class_outcome <- as.factor(data$class_outcome)
```

```{r}
numerical_cols <- sapply(data, is.numeric)
numerical_data <- data[, numerical_cols]

for (col in colnames(numerical_data)) {
  p <- ggplot(data, aes_string(x = col)) +
    geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
    ggtitle(paste("Distribution of", col)) +
    theme_minimal()
  print(p)
}
```
Split data
```{r}
set.seed(1204)

trainIndex <- createDataPartition(
  data$class_outcome,
  p = 0.7,
  list = FALSE,
  times = 1
)

trainData <- data[trainIndex, ]
testData  <- data[-trainIndex, ]
```
Build initial GLM Model
```{r}
initial_model <- glm(
  class_outcome ~ engagement + participation + attendance +
    class_duration + seating_group + teaching_method,
  data = trainData,
  family = binomial()
)

summary(initial_model)
```
Model Adequacy check

```{r}
plot(initial_model)
```
Cook Distance
```{r}
cooksd <- cooks.distance(initial_model)

influential <- as.numeric(
  names(cooksd)[(cooksd > (4 / length(cooksd)))]
)

cat("Influential observations:", influential, "\n")
```
Plot cook distance
```{r}
plot(cooksd,
 main = "Cook's Distance",
 xlab = "Observation Index",
 ylab = "Cook's Distance",
 type = "h", 
 col = ifelse(cooksd > (4 / length(cooksd)), "maroon", "purple")) 
abline(h = 4 / length(cooksd), col = "maroon", lty = 2)
influential_points <- which(cooksd > (4 / length(cooksd)))
text(influential_points,
 cooksd[influential_points],
 labels = influential_points,
 pos = 3,
 cex = 0.8,
 col = "maroon")
```
Refit model after removing influential points
```{r}
if (length(influential) > 0) {
  trainData <- trainData[-influential, ]
}

refit_model <- glm(
  class_outcome ~ engagement + participation + attendance +
    class_duration + seating_group + teaching_method,
  data = trainData,
  family = binomial()
)

summary(refit_model)
```
STEP WISE REGRESSION (AIC)
```{r}
final_model <- step(refit_model, direction = "both", trace = 1)
summary(final_model)
```
PLOT FINAL MODEL 

```{r}
plot(final_model)
```
EVALUATE THE FINAL MODEL USING AIC AND BIC

```{r}
aic_value <- AIC(final_model)
bic_value <- BIC(final_model)
cat("Final Model AIC: ", aic_value, "\n")
cat("Final Model BIC: ", bic_value, "\n")
```

```{r}
anova_comparison <- anova(refit_model, final_model, test = 'Chisq')
cat("Model Comparison (Chi-Square Test):\n")
print(anova_comparison)
```
PREDICT TEST SET USING FINAL MODEL

```{r}
testData$predicted <- predict(final_model, newdata = testData, type = "response")
predicted_classes <- ifelse(testData$predicted > 0.5, 1, 0)
confusionMatrix <- table(Predicted = predicted_classes, Actual = as.numeric(testData$class_outcome))
print(confusionMatrix)
```
ACCURACY CALCULATION
```{r}
accuracy <- sum(diag(confusionMatrix)) / sum(confusionMatrix)
cat("Final Model Accuracy: ", accuracy, "\n")
```
OVERDISPERSION AND MULTICOLLINEARITY
```{r}
residual_deviance <- deviance(final_model)
degrees_of_freedom <- df.residual(final_model)
overdispersion <- residual_deviance / degrees_of_freedom
cat("Residual Deviance: ", residual_deviance, "\n")
```

```{r}
cat("Degrees of Freedom: ", degrees_of_freedom, "\n")
```

```{r}
cat("Overdispersion Ratio: ", overdispersion, "\n")
```


```{r}
if (overdispersion > 1) {
 cat("Warning: Overdispersion detected in the model. Consider using a quasi-binomial or negative binomi
al approach.\n")
}
vif_values <- vif(final_model)
cat("Variance Inflation Factors (VIF):\n")
print(vif_values)
```


```{r}
if (any(vif_values > 5)) {
 cat("Warning: Multicollinearity detected. Consider removing or combining correlated predictors.\n")
}
```

